---
title: "primary_data"
author: "Jose Rico"
date: "8/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Importing Data

FiveThirtyEight's Polls Project <https://projects.fivethirtyeight.com/polls/> provides access to the most recent polls at  <https://github.com/fivethirtyeight/data>.  They provide a **url** for direct access the data in **csv** format

```{r data upload}
primary <- read.csv("https://projects.fivethirtyeight.com/polls-page/president_primary_polls.csv")
primary_t <- tibble(primary)
```

## Examine Structure of Data

If you want to quickly see data, just enter the data set's name, `primary` or click on it in the **Environment** tab

```{r}

```

Sometimes the original data frame in too large, you can use the `select()` command from **tidyverse**   
Note: 'Select' is for the columns of the data frame (aka the variables).

```{r}
primary_a <- primary %>%
     select(poll_id, state, pollster, fte_grade, sample_size, population, population_full, methodology, end_date, nationwide_batch, answer, pct) 
```

Now I only select the top 10 candidate for the next debate, similar code but this is a `filter` command from **tidyverse**
Note: 'Filter' is for the rows of the data frame (aka the observations).


```{r}
primary_a1 <- primary_a %>%
  rename(candidate_name=answer)

```
work on candidate_name data type (maybe that's the problem?)
```{r}

```


Filter out top candidates (using code from R4DS not moderndive)
```{r}
primary_a2 <- filter(primary_a1, candidate_name %in% c("Biden", "Gabbard", "Booker", "Castro", "Sanders", "Gillibrand", "Harris", "O'Rourke", "Buttigieg", "Warren", "Kolbuchar", "Yang", "Steyer"))
```

Filter out based on fte_grade A, A-, B+, B, B- keep
```{r}
primary_a3 <- filter(primary_a2, fte_grade %in% c("A", "A-", "B+", "B", "B-"))

```

Filter out just polls conducted either strictly  **live phone** or **online**  
```{r}
primary_a4 <- filter(primary_a3, methodology %in% c("Live Phone", "Online"))
```

View timeline of just a couple candidate polling at a time.

```{r}
primary_a5 <-filter(primary_a4, candidate_name %in% c("Sanders", "Biden" , "Harris", "Warren", "O'Rourke", "Buttigieg"))

```

Filter since just before the first debate
```{r}
primary_a6 <-filter(primary_a5, poll_id > 58488)
```


View the each candidate's percentage data from all polls in a visual summary know boxplot.
Choose 2 or 3 (2 is all polls, 3 is only A and B fte_rated polls)
```{r}
ggplot(data=primary_a6, mapping = aes(x=candidate_name, y=pct, color=candidate_name))+
     geom_boxplot()
```
Try to get the 2 on the same output
```{r}
ggplot(data=primary_a7, mapping = aes(x=candidate_name, y=pct, color=fte_grade))+
     geom_boxplot()
```
or view each grade on separate plot
```{r}
ggplot(data=primary_a7, mapping = aes(x=candidate_name, y=pct, color=candidate_name))+
    geom_boxplot()+
    coord_flip()+
    facet_wrap(~ methodology, nrow=2 )
```



Trendlines data from before first debate to know

```{r}
ggplot(data=primary_a7, mapping = aes(x=en_date1, y=pct, color=candidate_name ))+
    geom_point()+
    geom_smooth(method="lm", se = FALSE)
```
Prefer to use dates on x axis need to convert dates to days to election
first convert data format to `date`
```{r}
primary_a7 <- primary_a6 %>% 
  mutate(en_date1 = as.Date(end_date, format = "%m/%d/%Y"))

```
next create new variable days_to_NH
need **lubridate** package  (still working on this)
```{r}
NH_primary <- "2/11/19"
NH_primary <- as.Date(NH_primary, format = "%m/%d/%Y")
#primary_a8 <- primary_a7 %>% 
#  mutate(days_to_NH = ddays(NH_primary -en_date1 ))
```
Now, how do we subtract dates to creat the new column (easy once variable is in date format)
```{r}
primary_a8 <- primary_a7 %>% 
   mutate(days_to_NH = NH_primary -en_date1 )
```
Plot pct vs days_to_NH

```{r}
ggplot(data=primary_a8, mapping = aes(x=days_to_NH, y=pct, color=candidate_name ))+
    geom_point()+
    geom_smooth()

```




Summary Data on Each Candidate

```{r}
ggplot(data=filter(primary_a8, candidate_name %in% c("Sanders","Harris") & population == "lv" ), mapping = aes(x=days_to_NH, y=pct, color = candidate_name, size=sample_size ))+
    geom_point()+
    geom_smooth(method ="lm")
```




Histograms of each candidate poll numbers
```{r}
ggplot(data=primary_a5, mapping = aes(pct, color=candidate_name, size = sample_size ))+
    geom_histogram() +
    facet_wrap(~ candidate_name)
```



Select polls ending after June 18
```{r}
primary_a6 <-filter(primary_a5, poll_id > 58488)


```





